Death by Dropdown? Engineer Insightful Shiny Apps with Behavioral Science

### 2025-09-09 conference-materials/2025_posit-conf/draft.qmd
Revamp posit::conf 2025 slides with BID narrative & 16:9 layout
- Add conference context: subtitle `posit::conf(2025)`, Dublin Core date meta, visible date on title slide, and a persistent footer (“posit::conf(2025) | jrwinget.com”).
- Restructure talk into a clearer BID storyline:
  - New opener (“Dashboard Support Group”), “Why Listen to Me?”, and “The Problem: 18 Filters, 0 Insights”.
  - NOTICE/ANTICIPATE/STRUCTURE/VALIDATE sections with telemetry snippets, framing toggles, callouts, and a live demo Before → After.
  - Shorten the central question and reduce personas to a single “Product M

---
title: "Death by Dropdown? Engineer Insightful Shiny Apps with Behavioral Science"
author: "Jeremy Winget, PhD"
execute:
  echo: true
format:
  revealjs:
    theme: default
    slide-number: true
    transition: fade
---

## The Problem {.center background-color="#4682b4"}

::: {.large}
**Ever open a dashboard, see 18 filters across 6 tabs, and then...close your laptop?**
:::

::: {.fragment .large}
**Or just get the urge to...**

<div class="tenor-gif-embed" data-postid="17038619" data-share-method="host" data-aspect-ratio="1.77778" data-width="100%"><a href="https://tenor.com/view/ron-swanson-parks-and-recreation-parks-and-rec-computer-dumpster-gif-17038619">Ron Swanson Parks And Recreation GIF</a>from <a href="https://tenor.com/search/ron+swanson-gifs">Ron Swanson GIFs</a></div> <script type="text/javascript" async src="https://tenor.com/embed.js"></script>
:::

::: {.notes}
:::

## Quick Show of Hands 🙋 {.center background-color="#34495e"}

@@

::: {.fragment}
**Welcome to my impromptu Dashboard Support Group. You're among friends.** 😅
:::

::: {.notes}
- Confession time...
- Who's built a one of those dashboards? ...me
- Who's been forced to build one bc the client just kept asking for one more filter? ...me
- Who's ever used selectInput more than they probably should have? ...me
- You're not alone!
:::

## Why Me? {background-color="#2c3e50"}

:::: {.columns}
::: {.column width="60%"}
### Applied Social Psychologist → Full-Stack Engineer

- **PhD in Applied Psychology** studying how people process information
- **Led Shiny teams** at scale, providing enterprise dashboards
- **Core insight:** Most dashboard failures aren't technical, they're psychological
:::

::: {.column width="40%"}
::: {.fragment}
*"As a behavioral scientist in the tech world, I've seen firsthand how a dash of psychology can rescue users from that feeling of **'death by dropdown'"***
:::
:::
::::

## Your Dashboard = A Road Trip 🚗 {.center}

:::: {.columns}
::: {.column width="50%"}
### You wouldn't make a long drive without:
- **I**nterpreting travel guides
- **N**oticing any mechanical problems
- **A**nticipating roadblocks ahead
- **S**tructuring the best route
- **V**alidating everyone arrived happy
:::

::: {.column width="50%"}
::: {.fragment}
### So why build dashboards without:
- **I**nterpreting user needs
- **N**oticing friction points
- **A**nticipating behavior
- **S**tructuring a journey
- **V**alidating understanding
:::
:::
::::

::: {.notes}
- 20 seconds
- "This is the BID framework"
- Transition: "Let me show you the GPS"
:::

@@

**Remember - "Interfaces Need A Smart Vision"**

::: {.fragment}
**Behavioral Insight Design (BID)**: *Not just UX advice - it's behavioral science in your dev workflow*
:::

::: {.notes}
- 30-45 seconds MAX
- "Today we focus on the middle three - where psychology transforms dashboards"
- Quick transition to stages
:::

## BEFORE: A Familiar Nightmare

![](img/ui-before.png)

::: {.incremental}
- 18 filters • 6 tabs • no story • all data
- **User reaction:** "Where do I even start?"
- **Result:** 65% abandon rate in first 30 seconds
:::

## Every Dev's First Dashboard

![](img/this-is-fine.jpg)<br>
*"But it has ALL the features!"*

## INTERPRET: Plan Your Destination 🗺️ {background-color="#4682b4"}

```{r, interpret-stg}
library(dplyr)
library(bidux)

# Define the destination and goals
interpret_stage <- bid_interpret(
  central_question = "Which markets are driving performance, and where are we lagging?",
  data_story = list(
    hook = "Q4 revenue hit record high, but satisfaction dipped",
    context = "After aggressive marketing across all regions",
    tension = "West region satisfaction fell 10 points",
    resolution = "Focus retention efforts on underperforming regions"
  ),
  user_personas = list(
    list(
      name = "Product Managers",
      goals = "Monitor weekly KPIs",
      pain_points = "Too many filters to find important insights",
      technical_level = "Moderate"
    ),
    list(
      name = "Executives",
      goals = "Review quarterly trends",
      pain_points = "Overwhelmed by data, need clear takeaways",
      technical_level = "Basic"
    )
  )
)

@@

::: {.fragment}
**Key insight:** If you can't write the question in 10 words, users won't find answers in 100 filters
:::

::: {.notes}
- 1.5 minutes TOTAL for Interpret
- "You think you do this, but telemetry says otherwise"
- "Most jump straight to filters"
- Transition: "But here's what you're missing..."
:::

## NOTICE: The Warning Lights You're Ignoring ⚠️ {background-color="#e74c3c"}

@@

::: {.column width="50%"}
![](img/telemetry-sessions.png)
:::
::::

::: {.incremental}
- **Your assumption:** "They need all 18 filters"
- **Reality:** 82% use only 3 filters
- **Hidden cost:** 4.3 minutes to first click
:::

::: {.notes}
- START 3-minute NOTICE section
- "This is where behavioral science starts"
:::

## NOTICE: What Telemetry Reveals

```{r, telemetry-data}
#| echo: false

conn_telemetry_db <- function(db_path = "telemetry.sqlite") {
  # Create SQLite database with telemetry data
  con <- DBI::dbConnect(RSQLite::SQLite(), db_path)

  set.seed(456)
  n_sessions <- 1000

  # Create telemetry events with proper columns expected by bid_telemetry
  events <- data.frame()

  for(i in 1:n_sessions) {
    session_id <- paste0("session_", i)
    user_id <- paste0("user_", sample(1:200, 1))
    session_start <- Sys.time() - runif(1, 0, 30*24*60*60)  # Random time in last 30 days
    session_duration <- rexp(1, rate = 1/260)  # avg 4.3 minutes

    # Navigation events (page views)
    visited_tabs <- sample(
      c("Revenue Analysis", "Customer Insights", "Operations",
        "Product Performance", "Marketing Analytics", "Executive Summary"),
      sample(1:3, 1),
      prob = c(0.35, 0.25, 0.15, 0.10, 0.10, 0.05)
    )

    for(tab in visited_tabs) {
      events <- rbind(events, data.frame(
        timestamp = format(session_start + runif(1, 0, session_duration), "%Y-%m-%d %H:%M:%S"),
        session_id = session_id,
        user_id = user_id,
        event_type = "navigation",
        navigation_id = tab,
        input_id = NA,
        value = NA,
        error_message = NA,
        output_id = NA,
        stringsAsFactors = FALSE
      ))
    }

    # Input interaction events
    all_inputs <- c(
      "year", "quarter", "region",  # Used by 82%
      "rev_product", "rev_category", "rev_channel",
      "cust_segment", "cust_source", "cust_campaign",
      "ops_team", "ops_priority", "ops_status",
      "prod_product", "prod_compare",
      "mkt_campaign", "mkt_channel",
      "exec_metric", "exec_view"
    )

    if(runif(1) < 0.82) {
      # 82% use only first 3 filters
      used_inputs <- sample(all_inputs[1:3], sample(1:3, 1))
    } else {
      # 18% use additional filters
      n_filters <- sample(4:10, 1)
      used_inputs <- sample(all_inputs, n_filters)
    }

    for(input in used_inputs) {
      events <- rbind(events, data.frame(
        timestamp = format(session_start + runif(1, 0, session_duration), "%Y-%m-%d %H:%M:%S"),
        session_id = session_id,
        user_id = user_id,
        event_type = "input",
        navigation_id = NA,
        input_id = input,
        value = sample(c("Option1", "Option2", "All"), 1),
        error_message = NA,
        output_id = NA,
        stringsAsFactors = FALSE
      ))
    }

    # Error events (31% misinterpretation rate)
    if(runif(1) < 0.31) {
      error_messages <- c(
        "Filter combination produced no results",
        "Chart rendering failed",
        "Incorrect time period selected",
        "Data not available for selection"
      )

      events <- rbind(events, data.frame(
        timestamp = format(session_start + runif(1, 60, session_duration), "%Y-%m-%d %H:%M:%S"),
        session_id = session_id,
        user_id = user_id,
        event_type = "error",
        navigation_id = NA,
        input_id = NA,
        value = NA,
        error_message = sample(error_messages, 1),
        output_id = sample(c("revenuePlot", "satisfactionPlot", "revenueTable"), 1),
        stringsAsFactors = FALSE
      ))
    }

    # Login event at session start
    events <- rbind(events, data.frame(
      timestamp = format(session_start, "%Y-%m-%d %H:%M:%S"),
      session_id = session_id,
      user_id = user_id,
      event_type = "login",
      navigation_id = NA,
      input_id = NA,
      value = NA,
      error_message = NA,
      output_id = NA,
      stringsAsFactors = FALSE
    ))
  }

  # Sort events by timestamp
  events <- events[order(events$timestamp), ]

  # Write to SQLite database
  DBI::dbWriteTable(con, "events", events, overwrite = TRUE)

  DBI::dbDisconnect(con)

  return(db_path)
}
```

```{r, notice-stg}
# Using production telemetry
telemetry_path <- conn_telemetry_db("dashboard_telemetry.sqlite")

# bid_telemetry automatically categorizes and prioritizes issues!
issues <- bid_telemetry(telemetry_path)
head(issues)

# if we have telemetry, let bidux convert issues to Notice stages
critical_issues <- issues |>
  filter(severity == "critical") |>
  slice_head(n = 3)

notices <- bid_notices(
  issues = critical_issues,
  previous_stage = interpret_stage
)

notice_stage <- notices[[1]]

# Using observations/feedback
notice_stage <- bid_notice(
  previous_stage = interpret_stage,
  problem = "Users overwhelmed by 18 filters across 6 tabs",
  evidence = "82% of sessions use only 3 filters; 4.3 min to first insight"
  # NOTE: No theory specified - bidux will suggest one!
)

summary(notice_stage)
```

::: {.fragment}
**The "Just One More Filter" Death Spiral:**

![](img/anakin.jpg)
:::

::: {.notes}
- "Telemetry is your behavioral microscope"
- "Every hover tells a story of confusion"
- Transition: "Now let's talk about what happens in their heads"
:::

## NOTICE: Cognitive Load Measurement

:::: {.columns}
::: {.column width="50%"}
### You're Measuring:
- Click-through rates
- Time on page
- Completions
:::

::: {.column width="50%"}
### You Should Measure:
- **Decision fatigue points**
- **Paradox of choice moments**
- **Cognitive overflow indicators**
:::
::::

::: {.notes}
- "This is behavioral science in action"
- End NOTICE section (3 min total)
- Transition: "Once you spot problems, anticipate how users will cope"
:::

## ANTICIPATE: Pack for Cognitive Weather 🌦️ {background-color="#f39c12"}

@@

![](https://media.giphy.com/media/l0IykOsxLECVejOzm/giphy.gif)

*Users interpreting your carefully crafted visualizations*

::: {.fragment}
*And they'll write their own story if you don't guide them*
:::

::: {.notes}
- START 3-minute ANTICIPATE section
- "This is pure psychology"
- So we need to prepare for cognitive detours
:::

## ANTICIPATE: The Biases You're Not Defending Against

:::: {.columns}
::: {.column width="50%"}
### Anchoring Bias

```{r}
#| eval: false

# Bad: First number sticks
valueBox(value = -5, subtitle = "Change")

# Good: Context first
valueBox(value = "+15% YTD", subtitle = "Despite -5 this week")
```
:::

::: {.column width="50%"}
### Framing Effects

```{r}
# Don't specify bias_mitigations - let {bidux} suggest them!
anticipate_stage <- bid_anticipate(
  previous_stage = notice_stage,
  # bias_mitigations = NULL,
  include_accessibility = TRUE
)
```
:::
::::

::: {.notes}
- "Small changes, massive perception shifts"
- "This is what behavioral scientists know that devs don't"
:::

## ANTICIPATE: Confirmation Bias Protection

<!-- ![](img/confirmation-bias-toggle.png) -->

::: {.fragment}
**Result:** 74% reduction in misinterpretation
:::

::: {.notes}
- "Protect users from themselves"
- End ANTICIPATE (3 min total)
- Transition: "Now structure the journey"
:::

## STRUCTURE: The Scenic Route 🛣️ {background-color="#16a085"}

![](img/oprah.jpg)

::: {.fragment}
**If everything's important, nothing is.**
:::

::: {.notes}
- START 3.5-minute STRUCTURE section
- "This is where psychology meets code"
:::

## STRUCTURE: The Behavioral Science Hierarchy

```{r}
flags <- bid_flags(issues)

@@

structure_stage$suggestions
```

::: {.fragment}
### The F-Pattern Reading Path
![](img/f-pattern-heatmap.png)

*Place critical info where eyes naturally go*
:::

@@

## STRUCTURE: Smart Defaults from Telemetry

```{r}
all_suggestions <- c(
  interpret_stage$suggestions,
  notice_stage$suggestions,
  anticipate_stage$suggestions,
  structure_stage$suggestions
)

all_suggestions
```

::: {.fragment}
**Before:** User configures dashboard<br>
**After:** Dashboard anticipates user
:::

:::: {.columns}
::: {.column width="50%"}
### Before: Chaos
<!-- ![](img/ui-before.png) -->

- All options visible
- Competing for attention
- Analysis paralysis
:::

::: {.column width="50%"}
### After: Journey
<!-- ![](img/after-structured.png) -->

- Start with one question
- Reveal complexity gradually
- Cognitive scaffolding
:::
::::

::: {.notes}
- End STRUCTURE (3.5 min total)
- Transition: "Quick validation check"
:::

@@

::: {.notes}
- 1.5 minutes TOTAL for Validate
- "Devs know summaries, but miss the psychology"
- "Plain language = cognitive relief"
- Transition: "Let's see the impact"
:::

## The Transformation: Proven Results 📊

:::: {.columns}
::: {.column width="50%"}
### Before BID
- 18 filters, 6 tabs
- 4.3 min to insight
- 65% abandonment
- 31% misinterpret data
:::

::: {.column width="50%"}
### After BID
- 3 smart filters
- 45 sec to insight (-82%)
- 85% completion (+143%)
- 8% misinterpret (-74%)
:::
::::

::: {.fragment}
### One Team's Story:
*"We thought users wanted more options. Telemetry showed they wanted more guidance. BID framework + {bidux} = 224% increase in weekly active users"*
:::

::: {.notes}
- 1.5 minutes for results
- "This isn't theory - it's production data"
- Transition: "Here's how to start"
:::

## Your Roadmap Forward 🚀 {.center background-color="#2c3e50"}

@@

### This Week's Challenge:

1. **Install {bidux}:** `install.packages("bidux")`
2. **Run telemetry on ONE dashboard:** `bid_notice(telemetry = TRUE)`
3. **Share your before/after:** [@jrwinget](https://bsky.app/profile/jrwinget.bsky.social)

::: {.fragment}
**🎁 First 5 transformations get free 1:1 dashboard review!**

<!-- ![QR for contact](img/qr-combined.png) -->
:::

::: {.notes}
- 1 minute close
- "No more death by dropdown!"
- "Questions?"
- Thank organizers
:::

>>>>

---
title: "Death by Dropdown? Engineer Insightful Shiny Apps with Behavioral Science"
subtitle: "posit::conf(2025)"
author: "Jeremy Winget, PhD"
date: "2025-09-17"
execute:
  echo: true
format:
  revealjs:
    slide-number: true
    transition: fade
    footer: "posit::conf(2025) | jrwinget\\.com"
    preview-links: auto
    code-block-border-left: true
    highlight-style: github
    auto-animate: true
    chalkboard: false
    width: 1600
    height: 900
---

## {.center background-color="#2c3e50" auto-animate=true}

::: {.r-fit-text}
**Ever open a dashboard with 18 filters...**
:::

:::: {.fragment}
::: {.r-fit-text style="color: #e74c3c;"}
**...and immediately close your laptop?**
:::

<div class="tenor-gif-embed" data-postid="17038619" data-share-method="host" data-aspect-ratio="1.77778" data-width="60%"><a href="https://tenor.com/view/ron-swanson-parks-and-recreation-parks-and-rec-computer-dumpster-gif-17038619">Ron Swanson Parks And Recreation GIF</a></div> <script type="text/javascript" async src="https://tenor.com/embed.js"></script>
::::

::: {.notes}
- 15 seconds MAX
- "We've ALL been there"
- Transition: "Quick show of hands"
:::

## Dashboard Support Group 🙋 {.center background-color="#34495e" auto-animate=true}

@@

::: {.fragment}
### Welcome to my impromptu support group. You're among friends! 😅

<br>

👋 **I'm Jeremy! Behavioral psychologist turned engineer**<br>
*Think of this space as our little therapy session for dropdown overload*
:::

::: {.notes}
- 30 seconds
- Get laughs with "dashboard anonymous"
- "Hi Jeremy" callback
- Transition: "Let me properly introduce myself"
:::

## Why Listen to Me? {background-color="#2c3e50"}

:::: {.columns}
::: {.column width="60%"}
### From Lab to Production

![](img/hex-bidux.png){width="250" .absolute top=150 left=650}

- **PhD in Applied Psychology** <br>*Studied how individuals and groups process information and make decisions*

- **Led Shiny teams at scale** <br>*Built enterprise dashboards that folks actually use*

- **Created {bidux} package** <br>*Your behavioral scientist in the console*

::: {.fragment}
### What is Behavioral Science?
*The study of how people **actually** make decisions (spoiler: not rationally)*
:::
:::

::: {.column width="40%" .fragment}
### My Core Insight

> "Most dashboard failures aren't technical, they're psychological. We're building for computers, not humans."

**Today:** Transform your dashboards using the science of human behavior
:::
::::

::: {.notes}
- 45 seconds
- "Behavioral science = how humans actually work"
- "Not how we think they work"
- Transition: "Let me show you the framework"
:::

## Your Dashboard = A Road Trip 🚗 {background-color="#34495e"}

:::: {.columns .fragment}
::: {.column width="50%"}
### You wouldn't make a long drive without:
- 📍 **I**nterpreting travel guides
- ⚠️ **N**oticing any mechanical problems
- 🌦️ **A**nticipating roadblocks ahead
- 🗺️ **S**tructuring the best route
- ✅ **V**alidating everyone arrived happy
:::

::: {.column width="50%"}
### So why build dashboards without:
- 📍 **I**nterpreting user needs
- ⚠️ **N**oticing friction points
- 🌦️ **A**nticipating behavior
- 🗺️ **S**tructuring a journey
- ✅ **V**alidating understanding
:::

::: {.callout-tip}
### Helpful mnemonic device:
**"Interfaces Need A Smart Vision"**
:::
::::

::: {.notes}
- 20 seconds
- "This is the core idea of the BID framework"
- Transition: "Let me show you the GPS"
:::

@@

**Behavioral Insight Design (BID)**: *Not just UX, it's behavioral science in your workflow*

::: {.notes}
- 20 seconds MAX
- "BID framework = your GPS"
- "Today we focus on the middle three - where psychology transforms dashboards"
- "Let's take a look at a example from a client I recently worked with"
:::

## The Problem: 18 Filters, 0 Insights {background-image="img/ui-before.png" background-size="contain" background-position="right"}

::: {.fragment .absolute top=500 left=-40}
### We gathered telemetry on behavior:
- 😵 **65% abandon** in 30 seconds
- ⏱️ **4.3 minutes** to first click
- 🎯 **31% misinterpret** the data
- 📊 **82% use only 3** of 18 filters
:::

::: {.fragment .absolute top=700 left=-40}
**This is behavioral science *screaming* at us**
:::

::: {.notes}
- 45 seconds
- "Let me show you what users actually do"
- Transition to telemetry
:::

## INTERPRET: Plan Your Destination 🗺️ {background-color="#3498db" auto-animate=true}

```{r interpret-stg}
library(dplyr)
library(bidux)

# Define the destination and goals
interpret_stage <- bid_interpret(
  central_question = "Which markets are driving performance?",
  data_story = list(
    hook = "Q4 revenue hit record high, but satisfaction dipped",
    context = "After aggressive marketing across all regions",
    tension = "West region satisfaction fell 10 points",
    resolution = "Focus retention efforts on underperforming regions"
  ),
  user_personas = list(
    list(
      name = "Product Manager",
      goals = "Monitor weekly KPIs",
      pain_points = "Too many filters to find important insights",
      technical_level = "Moderate"
    )
  )
)

@@

::: {.fragment .callout-tip}
## The 10-Word Guideline:
**Can't explain your dashboard's purpose in 10 words?**

Users won't find answers in 100 filters
:::

::: {.notes}
- 30-45 seconds ONLY
- "A lot of devs think they do this, but telemetry says otherwise"
- "Most jump straight to filters"
- Transition: "But here's what they're missing..."
:::

## NOTICE: The Telemetry Bombshell 💣 {background-color="#e74c3c" auto-animate=true}

@@

::: {.column width="50%"}
```{r}
#| eval: false
#| code-line-numbers: "4-5,9"
# Actual production data
bid_telemetry("dashboard.sqlite") |>
  summary()
#> 82% use only 3 filters
#> 4.3 min to first insight
#> 65% abandonment rate

# Let bidux diagnose the problem
bid_notice(telemetry = TRUE)
#> Theory: Cognitive Load Theory
#> Severity: CRITICAL
```
:::
::::

::: {.fragment}
![](img/anakin.jpg){.absolute top=400 right=50 width="400"}
:::

::: {.notes}
- 2 minutes on NOTICE
- "This is your behavioral microscope"
- "Every hover tells a story"
:::

## NOTICE: What 1000 Sessions Reveal {background-color="#e74c3c"}

:::: {.columns}
::: {.column width="60%"}
![](img/telemetry-sessions.png)
:::

::: {.column width="40%"}
### The Hidden Costs
::: {.incremental}
- **Hover time:** 3.2 sec before clicking
- **Help icons:** 68% hover, 0% click
- **Filter #7:** Abandonment spike
- **Errors:** "No results" 319 times
:::

::: {.fragment}
**Your users are drowning in choices**
:::
:::
::::

::: {.notes}
- "Cognitive load is measurable"
- Transition: "Once you spot problems, anticipate how users will cope"
- Transition: "How do we do that?"
:::

## ANTICIPATE: Users Create Their Own Stories 🎭 {background-color="#f39c12"}

@@

<div class="tenor-gif-embed" data-postid="22483518" data-share-method="host" data-aspect-ratio="1.58416" data-width="50%"><a href="https://tenor.com/view/its-always-sunny-in-philadelphia-charlie-conspiracy-gif-22483518">Charlie Conspiracy GIF</a></div>

::: {.fragment}
**Without guidance, they'll find patterns that don't exist**
:::

::: {.notes}
- Start of 2 minutes on ANTICIPATE
- "This is pure psychology"
:::

## ANTICIPATE: Same Data, Different Story {background-color="#f39c12" auto-animate=true}

:::: {.columns}
::: {.column width="50%"}
### Marketing Sees Success 📈
<!-- ![Placeholder: Dashboard showing "75% to Goal" with green progress bar](img/framing-positive-placeholder.png) -->

*"We're crushing it! Almost there!"*
:::

::: {.column width="50%"}
### Finance Sees Risk 📉
<!-- ![Placeholder: Same dashboard showing "25% Gap Remaining" with red deficit bar](img/framing-negative-placeholder.png) -->

*"We're behind! Need immediate action!"*
:::
::::

::: {.fragment}
```{r}
#| eval: false
# One toggle, two perspectives
bid_anticipate() |>
  add_framing_toggle(
    positive = "75% to goal",
    negative = "25% gap remaining"
  )
```

**Result: 74% reduction in misinterpretation**
:::

::: {.notes}
- "Same data, opposite conclusions"
- "Framing bias in action"
- Transition: "Now...let's structure the experience"
:::

## STRUCTURE: Stop the Filter Madness 🛑 {background-color="#16a085"}

![](img/oprah.jpg){.absolute top=100 right=50 width="500"}

:::: {.columns}
::: {.column width="50%"}
### Progressive Disclosure FTW

```{r}
#| eval: false
#| code-line-numbers: "2-4,6-7"
bid_structure() |>
  set_smart_defaults(based_on = "telemetry") |>
  limit_initial_choices(n = 3) |>
  defer_complexity()

#> Smart defaults from 1000 sessions:
#> Date: Last 30 days (87% usage)
#> Product: All (72% default)
#> Region: All (69% default)
```
<br>

```{r}
#| eval: false
flags <- bid_flags(issues)

@@

structure_stage$suggestions
```
:::
::::

::: {.notes}
- 2 minutes on STRUCTURE
- "Let the F-pattern guide you"
:::

## STRUCTURE: The F-Pattern Is Your Friend {background-color="#16a085" auto-animate=true}

![](img/f-pattern-heatmap.png){width="80%"}

*Place critical info where eyes naturally go*

@@

@@

::: {.notes}
- 20 minutes TOTAL for Validate
- "Devs know summaries, but miss the psychology"
- "Plain language = cognitive relief"
- Transition: "Let's see the impact"
:::

## The Transformation: Measurable Impact 📊 {auto-animate=true}

![](img/summary-report.png){width="90%"}

::: {.notes}
- 1 minute
- "This isn't theory - it's production data"
:::

## Live Demo: Before → After {.center background-color="#2c3e50"}

:::: {.columns}
::: {.column width="50%"}
### Before: Death by Dropdown
```{r}
#| eval: false
# 1_init.R - The nightmare
shinyApp(
  ui = fluidPage(
    sidebarPanel(
      selectInput("year", choices = years),
      selectInput("quarter", choices = quarters),
      selectInput("region", choices = regions),
      # ... 15 more filters
    )
  )
)
```
![](img/ui-before.png)
:::

::: {.column width="50%" .fragment}
### After: Guided Journey
```{r}
#| eval: false
# 3_redesign.R - BID applied
shinyApp(
  ui = page_navbar(
    card(
      card_header("Executive Summary"),
      uiOutput("executiveSummary")),
    # Only 3 smart filters
    # Progressive disclosure
  )
)
```
![](img/ui-after.png)
:::
::::

::: {.notes}
- 30 seconds
- "Let me show you the actual apps"
- Quick toggle between windows
:::

## Your Implementation Roadmap 🚀 {.center background-color="#27ae60"}

![](img/bid-framework.png){width="75%"}

@@

1. **Install {bidux}:** `install.packages("bidux")`
2. **Implement the BID framework** on one of your dashboards
3. **Share your before/after:** [@jrwinget](https://bsky.app/profile/jrwinget.bsky.social)
4. **Celebrate the end of death by dropdown!** 🎉

::: {.notes}
- 1 minute close
- "No more death by dropdown!"
- Transition: "And with that...I want to thank you all for being here"
:::

## Thank You! {.center background-color="#2c3e50"}

:::: {.columns}
::: {.column width="33%"}
![](img/hex-bidux.png)
:::

::: {.column width="33%"}
### 📚 Resources
- [github.com/jrwinget/bidux](https://github.com/jrwinget/bidux)
- [github.com/jrwinget/bid-framework](https://github.com/jrwinget/bid-framework)
:::

::: {.column width="33%"}
### 🤝 Connect
- <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-bluesky" viewBox="0 0 16 16">
  <path d="M3.468 1.948C5.303 3.325 7.276 6.118 8 7.616c.725-1.498 2.698-4.29 4.532-5.668C13.855.955 16 .186 16 2.632c0 .489-.28 4.105-.444 4.692-.572 2.04-2.653 2.561-4.504 2.246 3.236.551 4.06 2.375 2.281 4.2-3.376 3.464-4.852-.87-5.23-1.98-.07-.204-.103-.3-.103-.218 0-.081-.033.014-.102.218-.379 1.11-1.855 5.444-5.231 1.98-1.778-1.825-.955-3.65 2.28-4.2-1.85.315-3.932-.205-4.503-2.246C.28 6.737 0 3.12 0 2.632 0 .186 2.145.955 3.468 1.948"/>
</svg> [@jrwinget](https://bsky.app/profile/jrwinget.bsky.social)
- <svg xmlns="http://www.w3.org/2000/svg" width="16" height="16" fill="currentColor" class="bi bi-linkedin" viewBox="0 0 16 16">
  <path d="M0 1.146C0 .513.526 0 1.175 0h13.65C15.474 0 16 .513 16 1.146v13.708c0 .633-.526 1.146-1.175 1.146H1.175C.526 16 0 15.487 0 14.854zm4.943 12.248V6.169H2.542v7.225zm-1.2-8.212c.837 0 1.358-.554 1.358-1.248-.015-.709-.52-1.248-1.342-1.248S2.4 3.226 2.4 3.934c0 .694.521 1.248 1.327 1.248zm4.908 8.212V9.359c0-.216.016-.432.08-.586.173-.431.568-.878 1.232-.878.869 0 1.216.662 1.216 1.634v3.865h2.401V9.25c0-2.22-1.184-3.252-2.764-3.252-1.274 0-1.845.7-2.165 1.193v.025h-.016l.016-.025V6.169h-2.4c.03.678 0 7.225 0 7.225z"/>
</svg> [@jrwinget](https://www.linkedin.com/in/jrwinget/)
:::
::::

<br>

::: {.fragment}
### Special Thanks 🙏
*To the posit::conf organizers for making this possible*<br>
*To the Shiny team for the amazing framework*<br>
*To you for caring about your users*
:::

::: {.notes}
- "Thank you!"
- "Questions?"
:::
